---
title: "Cross validation 2008"
---

```
## 0: not-erg before and after (no gain)
## 1: not-erg before, erg after (gain)
## 2: erg before, not-erg after (loss)
## 3: erg before and after (no loss)
```

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse) ## load filter(), ... function
library(blockCV)
library(sf)
library(raster)
library(ggplot2)
library(shiny) ## what for ... ??
library(ROSE)
library(caret)
source("functions.R")
library(randomForest)
library(rpart.plot)
library(spdep)
library(rfPermute)
library(metRology)
library(Metrics)
library(vip)
```

To knit "by hand", `rmarkdown::render("Cross_validation.Rmd")`

```{r input_data}
## increase memory
if (.Platform$OS.type=="windows") memory.limit(1000000)
load("rr_points14.RData")
dat2008 <- dplyr::filter(rr_points14[["2008"]],
                     x<604000 & y >284000 &  y < 2846000)
dat2008$change2 <- change_fun(dat2008$change)
dat2008_gain <- dplyr::filter(dat2008, change2 %in% c("no gain", "gain")) 
dat2008_upsample <- ROSE::ovun.sample(change ~ ., data = dat2008_gain, method = "both", N=1500)$data

```


buffer based on the defined boders=buffer 3
```{r determine_data}
##introsucing the specific block 
corners2008 <- list(x=c(601500,605000),
                y=c(2840000,2836000))
pts2008 <- seq(nrow(dat2008_upsample))
##introducing test data in the introduced block
test2008 <- pts2008[dat2008_upsample$x > corners2008$x[1] &
            dat2008_upsample$x < corners2008$x[2] &
            dat2008_upsample$y < corners2008$y[1] &
            dat2008_upsample$y > corners2008$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train2008 <- setdiff(pts2008,test2008)
plot(y~x,data=dat2008_upsample[train2008,])
grid(nx=4,ny=4,col=5)
with(dat2008_upsample[test2008,], points(x,y,col=2,pch=16))
```
```{r classified_maps}
Classified1=raster("Classified1.tif")
Classified2=raster("classified2.tif")
```
The appropriate format of species data for the blockCV package is simple features (sf) or SpatialPointsDataFrame.pamake a SpatialPointsDataFrame object from data.frame
```{r numeric_data}
PA_data <- st_as_sf(dat2008_upsample, coords = c("x", "y"), crs = crs(Classified1))
```

determine train and test data 
```{r train_testdata}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train20082 <- PA_data[train2008,]
test20082 <- PA_data[test2008, ]
train20082$geometry <- NULL
test20082$geometry <- NULL
```
Random forest
```{r random_forest_buffer3}
rfb32008 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train20082, n.trees=250,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds)
plot(rfb32008)
varImpPlot(rfb32008)
predb32008 <- predict(rfb32008, newdata=test20082, typeprobs="prob", predict.all=TRUE)
conb2008 <- caret::confusionMatrix(predb32008$aggregate, as.factor(test20082$change2))
table2008=table(test20082$change2, predb32008$aggregate)
set_tree2008 = rpart(change2 ~ . -change -landuse ,
                 data =train20082)
rpart.plot(set_tree2008)
```
H-P:error:R2: NA Rsquare:R2
```{r Rsquare:R2}
accuracy2008=postResample(pred = predb32008$aggregate,obs = as.factor(test20082$change2))
print(accuracy2008)
print(paste0('R2: ' ,caret::postResample(pred = predb32008$aggregate,obs = as.factor(test20082$change2)['Rsquared'] )))
```
ROC
```{r roc_random_forest_buffer3_roc, message=TRUE}
ff2008 <-  factor(test20082$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
table(test20082$change)
table(ff2008)
## convert a two-level factor back to 0/1
to_binary2008 <- function(x) { as.numeric(x) -1 }
rocCurve2008 <- pROC::roc(to_binary2008(ff2008),
                      ## predb3[,"change"]
                      to_binary2008(predb32008$aggregate))
caret::confusionMatrix(ff2008, predb32008$aggregate)
## gain/no gain  vs.  change/no change
plot(rocCurve2008)

## undebug(pROC:::roc.default)
pp <- pROC::roc(sample(0:1,size=50,replace=TRUE), pred=runif(50))
plot(pp)
plot(pROC::roc(c(1,1,0,0), pred=c(0.2,0.4,0.1,0.2)))
plot(pROC::roc(c(1,0), pred=c(0.2,0.2)))
## we know this won't work because it only has one data point,
## we need at least one observation in each level
try(plot(pROC::roc(factor(1, levels=0:1), pred=0.2)), silent=TRUE)
 
plot(rocCurve2008 ,col=c(4))
###calculate the area under curve (bigger is better)
## extract AUC and drop extra information
auc <- function(x) c(x$auc)
a2008=auc(rocCurve2008)
```
The basic idea is to consider a variable important if it has a positive effect on the prediction accuracy (classification)
importance plots
```{r variable_importance_plot}
vi(rfb32008)
vip(rfb32008)
```
