---
title: "Cross validation 1997"
---

```
## 0: not-erg before and after (no gain)
## 1: not-erg before, erg after (gain)
## 2: erg before, not-erg after (loss)
## 3: erg before and after (no loss)
```
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse) ## load filter(), ... function
library(blockCV)
library(sf)
library(raster)
library(ggplot2)
library(shiny) ## what for ... ??
library(ROSE)
library(caret)
source("functions.R")
library(randomForest)
library(rpart.plot)
library(spdep)
library(rfPermute)
library(metRology)
library(Metrics)
library(vip)
```

To knit "by hand", `rmarkdown::render("Cross_validation.Rmd")`

```{r input_data}
## increase memory
if (.Platform$OS.type=="windows") memory.limit(1000000)
load("rr_points14.RData")
dat1997 <- dplyr::filter(rr_points14[["1997"]],
                     x<604000 & y >284000 &  y < 2846000)
dat1997$change2 <- change_fun(dat1997$change)
dat1997_gain <- dplyr::filter(dat1997, change2 %in% c("no gain", "gain")) 
dat1997_upsample <- ROSE::ovun.sample(change ~ ., data = dat1997_gain, method = "both", N=1500)$data

```

buffer based on the defined boders=buffer 3
```{r determine_data}
##introsucing the specific block 
corners1997 <- list(x=c(601500,605000),
                y=c(2840000,2836000))
pts1997 <- seq(nrow(dat1997_upsample))
##introducing test data in the introduced block
test1997 <- pts1997[dat1997_upsample$x > corners1997$x[1] &
            dat1997_upsample$x < corners1997$x[2] &
            dat1997_upsample$y < corners1997$y[1] &
            dat1997_upsample$y > corners1997$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train1997 <- setdiff(pts1997,test1997)
plot(y~x,data=dat1997_upsample[train1997,])
grid(nx=4,ny=4,col=5)
with(dat1997_upsample[test1997,], points(x,y,col=2,pch=16))
```

```{r classified_maps}
Classified1=raster("Classified1.tif")
Classified2=raster("classified2.tif")
```
The appropriate format of species data for the blockCV package is simple features (sf) or SpatialPointsDataFrame.pamake a SpatialPointsDataFrame object from data.frame
```{r numeric_data}
PA_data <- st_as_sf(dat1997_upsample, coords = c("x", "y"), crs = crs(Classified1))
```



determine train and test data 
```{r train_testdata}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train19972 <- PA_data[train1997,]
test19972 <- PA_data[test1997, ]
train19972$geometry <- NULL
test19972$geometry <- NULL
```
Random forest
```{r random_forest_buffer3}
rfb31997 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train19972, n.trees=250,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds)
plot(rfb31997)
varImpPlot(rfb31997)
predb31997 <- predict(rfb31997, newdata=test19972, typeprobs="prob", predict.all=TRUE)
conb1997 <- caret::confusionMatrix(predb31997$aggregate, as.factor(test19972$change2))
table1997=table(test19972$change2, predb31997$aggregate)

set_tree2 = rpart(change2 ~ . -change -landuse ,
                 data =train19972)
rpart.plot(set_tree2)
```
H-P:error:R2: NA Rsquare:R2
```{r Rsquare:R2}
accuracy1997=postResample(pred = predb31997$aggregate,obs = as.factor(test19972$change2))
print(accuracy1997)
print(paste0('R2: ' ,caret::postResample(pred = predb31997$aggregate,obs = as.factor(test19972$change2)['Rsquared'] )))
```
ROC
```{r roc_random_forest_buffer3_roc, message=TRUE}
ff1997 <-  factor(test19972$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
table(test19972$change)
table(ff1997)
## convert a two-level factor back to 0/1
to_binary1997 <- function(x) { as.numeric(x) -1 }
rocCurve1997 <- pROC::roc(to_binary1997(ff1997),
                      ## predb3[,"change"]
                      to_binary1997(predb31997$aggregate))
caret::confusionMatrix(ff1997, predb31997$aggregate)
## gain/no gain  vs.  change/no change
plot(rocCurve1997)

## undebug(pROC:::roc.default)
pp <- pROC::roc(sample(0:1,size=50,replace=TRUE), pred=runif(50))
plot(pp)
plot(pROC::roc(c(1,1,0,0), pred=c(0.2,0.4,0.1,0.2)))
plot(pROC::roc(c(1,0), pred=c(0.2,0.2)))
## we know this won't work because it only has one data point,
## we need at least one observation in each level
try(plot(pROC::roc(factor(1, levels=0:1), pred=0.2)), silent=TRUE)
 
plot(rocCurve1997 ,col=c(4))
###calculate the area under curve (bigger is better)
## extract AUC and drop extra information
auc <- function(x) c(x$auc)
a1997=auc(rocCurve1997)
```
The basic idea is to consider a variable important if it has a positive effect on the prediction accuracy (classification)
importance plots
```{r variable_importance_plot}
vi(rfb31997)
vip(rfb31997)
```
