---
title: "averaging AUC across folds"
---

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse) ## load filter(), ... function
library(blockCV)
library(sf)
library(raster)
library(ggplot2)
library(shiny) ## what for ... ??
library(ROSE)
library(caret)
source("functions.R")
library(randomForest)
library(rpart.plot)
library(spdep)
library(rfPermute)
library(metRology)
library(caTools)
library(sf)
library(rgeos)
library(readxl)
library(raster)
library(tidyverse) ## includes purrr, readr, dplyr, ...
library(fasterize)  ## may need to install this
library(rasterVis)
library(foreign)
library(cowplot)
library(lulcc)
library(sp)
library(broom)
library(ggplot2); theme_set(theme_bw())
library(dotwhisker)
library(future)
library(furrr)
library(bbmle)
library(DHARMa)
library(ResourceSelection)
library(arm)
library(brglm2)
library(SparseM)
library(Hmisc)
library(rms)

```

To knit "by hand", `rmarkdown::render("Cross_validation.Rmd")`

```{r input_data}
## increase memory
if (.Platform$OS.type=="windows") memory.limit(1000000)

## get all data (as tables of x,y, predictors, change)
load("rr_points14.RData")
## need
## R CMD BATCH --vanilla climate.R
## R CMD BATCH --vanilla winddir.R
## R CMD BATCH --vanilla basic.R

## test with smaller data set
dat <- dplyr::filter(rr_points14[["2014"]],
                     x<604000 & y >284000 &  y < 2846000)
dat$change2 <- change_fun(dat$change)

```

```{r upsample}
dat_gain <- dplyr::filter(dat, change2 %in% c("no gain", "gain")) 

## analyzing gain only, first ...
if (file.exists("dat_upsample.rda")) {
    load("dat_upsample.rda")
}  else {
    set.seed(101)
    dat_upsample <- ROSE::ovun.sample(change ~ ., data = dat_gain, method = "both", N=1500)$data
save(dat_upsample, file="dat_upsample.rda")
}

```

```{r classified_maps}
Classified1=raster("Classified1.tif")
Classified2=raster("classified2.tif")
```
The appropriate format of species data for the blockCV package is simple features (sf) or SpatialPointsDataFrame.pamake a SpatialPointsDataFrame object from data.frame
```{r numeric_data}
PA_data <- st_as_sf(dat_upsample, coords = c("x", "y"), crs = crs(Classified1))
```

```{r train_vs_test}
set.seed(23489)
train_index <- sample(1:nrow(PA_data), 0.9 * nrow(dat_upsample))
sample_train <- PA_data[train_index, ]
sample_test  <- PA_data[-train_index, ]
```
main block
```{r determine_data}
##introsucing the specific block 
corners <- list(x=c(601500,605000),
                y=c(2839900,2835800))
pts <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test <- pts[dat_upsample$x > corners$x[1] &
            dat_upsample$x < corners$x[2] &
            dat_upsample$y < corners$y[1] &
            dat_upsample$y > corners$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train <- setdiff(pts,test)
plot(y~x,data=dat_upsample[train,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test,], points(x,y,col=2,pch=16))
```
determine train and test data 
```{r train_testdata}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train2 <- PA_data[train,]
test2 <- PA_data[test, ]
train2$geometry <- NULL
test2$geometry <- NULL
```

```{r random_forest_buffer3}
rfb3 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train2, n.trees=250,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds, keep.inbag = TRUE)
```

```{r roc_random_forest_buffer3_roc, message=TRUE}
predb3 <- predict(rfb3, newdata=test2, typeprobs="prob", predict.all=TRUE)
ff <-  factor(test2$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
table(test2$change)
table(ff)
## convert a two-level factor back to 0/1
to_binary <- function(x) { as.numeric(x) -1 }
rocCurve <- pROC::roc(to_binary(ff),
                      ## predb3[,"change"]
                      to_binary(predb3$aggregate))
caret::confusionMatrix(ff, predb3$aggregate)
auc <- function(x) c(x$auc)
a1=auc(rocCurve)
```
block 1
```{r}
set.seed(101)
## can we find the corners of block 1 automatically?

corn1 <- list(x=c(601500,605000),
                y=c(2836000,2833500))
pts1 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test1 <- pts1[dat_upsample$x > corn1$x[1] &
            dat_upsample$x < corn1$x[2] &
            dat_upsample$y < corn1$y[1] &
            dat_upsample$y > corn1$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train1 <- setdiff(pts1,test1)
plot(y~x,data=dat_upsample[train1,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test1,], points(x,y,col=2,pch=16))
##no gain data
```
block2
```{r}
corn2 <- list(x=c(601500,605000),
                y=c(2843000,2839600))
pts2 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test2 <- pts2[dat_upsample$x > corn2$x[1] &
            dat_upsample$x < corn2$x[2] &
            dat_upsample$y < corn2$y[1] &
            dat_upsample$y > corn2$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train2 <- setdiff(pts2,test2)
plot(y~x,data=dat_upsample[train2,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test2,], points(x,y,col=2,pch=16))
```

```{r train_testdata_block2}
numFolds <- trainControl(method = "cv", number = 10)
train22 <- PA_data[train2,]
test22 <- PA_data[test2, ]
train22$geometry <- NULL
test22$geometry <- NULL
```

```{r random_forest_block2}
rfbc2 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train22, n.trees=400,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds, keep.inbag = TRUE)
plot(rfbc2)
varImpPlot(rfbc2)
predbc2 <- predict(rfbc2, newdata=test22, typeprobs="prob", predict.all=TRUE)
conb2 <- caret::confusionMatrix(predbc2$aggregate, as.factor(test22$change2))
```
ROC_random_forest_buffer
```{r roc_block2, message=TRUE}
ff2 <-  factor(test22$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
table(test22$change)
table(ff2)
## convert a two-level factor back to 0/1
to_binary <- function(x) { as.numeric(x) -1 }
rocCurve2 <- pROC::roc(to_binary(ff2),
                      ## predb3[,"change"]
                      to_binary(predbc2$aggregate))
plot(rocCurve2)
auc <- function(x) c(x$auc)
a2=auc(rocCurve2)
```
block3
```{r}
corn3 <- list(x=c(601500,605000),
                y=c(28465000,2843000))
pts3 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test3 <- pts3[dat_upsample$x > corn3$x[1] &
            dat_upsample$x < corn3$x[2] &
            dat_upsample$y < corn3$y[1] &
            dat_upsample$y > corn3$y[2]]

train3 <- setdiff(pts3,test3)
plot(y~x,data=dat_upsample[train3,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test3,], points(x,y,col=2,pch=16))
##no gain
```
block4
```{r}
corn4 <- list(x=c(598500,601700),
                y=c(2843100,2839500))
pts4 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test4 <- pts4[dat_upsample$x > corn4$x[1] &
            dat_upsample$x < corn4$x[2] &
            dat_upsample$y < corn4$y[1] &
            dat_upsample$y > corn4$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train4 <- setdiff(pts4,test4)
plot(y~x,data=dat_upsample[train4,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test4,], points(x,y,col=2,pch=16))
```

```{r train_testdata_block4}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train44 <- PA_data[train4,]
test44 <- PA_data[test4, ]
train44$geometry <- NULL
test44$geometry <- NULL
```
Random forest
```{r random_forest_block4}
rfbc4 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train44, n.trees=300,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds, keep.inbag = TRUE)
plot(rfbc4)
varImpPlot(rfbc4)
predbc4 <- predict(rfbc4, newdata=test44, typeprobs="prob", predict.all=TRUE)
conb4 <- caret::confusionMatrix(predbc4$aggregate, as.factor(test44$change2))
```

```{r roc_block4, message=TRUE}
ff4 <-  factor(test44$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
to_binary4 <- function(x) { as.numeric(x) -1 }
rocCurve4 <- pROC::roc(to_binary(ff4),
                      ## predb3[,"change"]
                      to_binary(predbc4$aggregate))
plot(rocCurve4)
auc <- function(x) c(x$auc)
a3=auc(rocCurve4)
```
block5
```{r}
corn5 <- list(x=c(598500,601500),
                y=c(2839500,2836000))
pts5 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test5 <- pts5[dat_upsample$x > corn5$x[1] &
            dat_upsample$x < corn5$x[2] &
            dat_upsample$y < corn5$y[1] &
            dat_upsample$y > corn5$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train5 <- setdiff(pts5,test5)
plot(y~x,data=dat_upsample[train5,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test5,], points(x,y,col=2,pch=16))
##no gain
```
block6
```{r}
corn6 <- list(x=c(596000,598500),
                y=c(28465000,2842800))
pts6 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test6 <- pts6[dat_upsample$x > corn6$x[1] &
            dat_upsample$x < corn6$x[2] &
            dat_upsample$y < corn6$y[1] &
            dat_upsample$y > corn6$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train6 <- setdiff(pts6,test6)
plot(y~x,data=dat_upsample[train6,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test6,], points(x,y,col=2,pch=16))
```

```{r train_testdata_block6}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train66 <- PA_data[train6,]
test66 <- PA_data[test6, ]
train66$geometry <- NULL
test66$geometry <- NULL
```
Random forest
```{r random_forest_block6}
rfbc6 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train66, n.trees=350,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds, keep.inbag = TRUE)
plot(rfbc4)
varImpPlot(rfbc6)
predbc6 <- predict(rfbc6, newdata=test66, typeprobs="prob", predict.all=TRUE)
conb6 <- caret::confusionMatrix(predbc6$aggregate, as.factor(test66$change2))
```

```{r roc_block6, message=TRUE}
ff6 <-  factor(test66$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
to_binary6 <- function(x) { as.numeric(x) -1 }
rocCurve6 <- pROC::roc(to_binary(ff6),
                      ## predb3[,"change"]
                      to_binary(predbc6$aggregate))
plot(rocCurve6)
auc <- function(x) c(x$auc)
a4=auc(rocCurve6)
```
block7
```{r}
corn7 <- list(x=c(595500,598500),
                y=c(2843500,2835000))
pts7 <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test7 <- pts7[dat_upsample$x > corn7$x[1] &
            dat_upsample$x < corn7$x[2] &
            dat_upsample$y < corn7$y[1] &
            dat_upsample$y > corn7$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train7 <- setdiff(pts7,test7)
plot(y~x,data=dat_upsample[train7,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test7,], points(x,y,col=2,pch=16))
```

```{r train_testdata_block7}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train77 <- PA_data[train7,]
test77 <- PA_data[test7, ]
train77$geometry <- NULL
test77$geometry <- NULL
```
Random forest
```{r random_forest_block7}
rfbc7 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train77, n.trees=350,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds, keep.inbag = TRUE)
plot(rfbc7)
varImpPlot(rfbc7)
predbc7 <- predict(rfbc7, newdata=test77, typeprobs="prob", predict.all=TRUE)
conb7 <- caret::confusionMatrix(predbc7$aggregate, as.factor(test77$change2))
```

```{r roc_block7, message=TRUE}
ff7 <-  factor(test77$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
to_binary7 <- function(x) { as.numeric(x) -1 }
rocCurve7 <- pROC::roc(to_binary(ff7),
                      ## predb3[,"change"]
                      to_binary(predbc7$aggregate))
plot(rocCurve7)
auc <- function(x) c(x$auc)
a5=auc(rocCurve7)
```

blpck8
```{r determine_data_2}
##introsucing the specific block 
corner8 <- list(x=c(593500,595500),
                y=c(28465000,2843000))
pts <- seq(nrow(dat_upsample))
##introducing test data in the introduced block
test8 <- pts[dat_upsample$x > corner8$x[1] &
            dat_upsample$x < corner8$x[2] &
            dat_upsample$y < corner8$y[1] &
            dat_upsample$y > corner8$y[2]]

##setdiff indicates which elements of a vector or data frame X are not existent in a vector or data frame Y.
train8 <- setdiff(pts,test8)
plot(y~x,data=dat_upsample[train8,])
grid(nx=4,ny=4,col=5)
with(dat_upsample[test8,], points(x,y,col=2,pch=16))
```

```{r train_testdata_block8}
##number of k
numFolds <- trainControl(method = "cv", number = 10)
##trainand
train88 <- PA_data[train8,]
test88 <- PA_data[test8, ]
train88$geometry <- NULL
test88$geometry <- NULL
```
Random forest
```{r random_forest_block8}
rfbc8 <- randomForest(formula= factor(change2) ~ . - landuse - change,
                       data = train88, n.trees=350,interaction.depth=7,
                       ## do.trace=1,
                       type="classification", proximity=TRUE, trcontrol=numFolds, keep.inbag = TRUE)
plot(rfbc8)
varImpPlot(rfbc8)
predbc8 <- predict(rfbc8, newdata=test88, typeprobs="prob", predict.all=TRUE)
conb8 <- caret::confusionMatrix(predbc8$aggregate, as.factor(test88$change2))
```

```{r roc_block8, message=TRUE}
ff8 <-  factor(test88$change,
              ## 0 = no gain,  1 = gain
              levels=c(0,1),  ## these are the values in the original vector
              labels=c("no gain","gain") ## these will be the names in the new vector
              )
to_binary8 <- function(x) { as.numeric(x) -1 }
rocCurve8 <- pROC::roc(to_binary(ff8),
                      ## predb3[,"change"]
                      to_binary(predbc8$aggregate))
plot(rocCurve7)
auc <- function(x) c(x$auc)
a6=auc(rocCurve8)
```

```{r}
x=c(a1,a2,a3,a4,a5,a6)
y=mean(x)
print(y)
```
Out of bag (OOB) score is a way of validating the Random forest model.
```{r Out_of_Bag_error}
err1=rfb3$err.rate
oob1=err[nrow(err),"OOB"]
err2=rfbc2$err.rate
oob2=err[nrow(err2),"OOB"]
err3=rfbc4$err.rate
oob3=err[nrow(err3),"OOB"]
err4=rfbc6$err.rate
oob4=err[nrow(err4),"OOB"]
err5=rfbc7$err.rate
oob5=err[nrow(err5),"OOB"]
err6=rfbc6$err.rate
oob6=err[nrow(err6),"OOB"]
x2=c(oob1,oob2,oob3,oob4,oob5,oob6)
y2=mean(x2)
print(y2)
x3=c(rfb3,rfbc2,rfbc4,rfbc6,rfbc7,rfbc8)
```