#extra codes
##random forest
```{r}
library(randomForest)
library(caTools)
library(caret)
library(ranger)
library(ROSE)
```
let's just select the western tip of the study areaget 2014 data
```{r}
testdat <- filter(rr_points14[["2014"]],
                  x<604000 & y >284000 &  y < 2846000)
```
upsample & downsample
```{r}
table(testdat$change)
testdat$change <- factor(testdat$change, levels=0:3,
                         labels= c("no gain","gain","loss","no loss"))
testdat$change2 <- factor(ifelse(testdat$change %in% c("no gain", "no loss"),
                                 "no change", "change"))
prop.table(x=table(a2$change))
prop.table(x=table(a2$change2))
testdatss <- ovun.sample(change2~., data = testdat, method = "both", N=1500)$data
ss2 <- ovun.sample(change2~., data = a2, method = "under")$data
```
generate a sequence of random numbers – it ensures that you get the same result if you start with that same seed each time you run the same process.
```{r}
set.seed(123)
sample <- sample.split(ss2$change2, SplitRatio = 0.75)
```
subset=take random samples from a dataset
REGRESSION: trying to predict {0,1,2,3} [change values]this doesn't make sense because
*0: not-erg before and after (no gain)
*1: not-erg before, erg after (gain)
*2: erg before, not-erg after (loss)
*3: erg before and after (no loss)
regression might make sense if we treated gain {0,1} and loss {2,3}separately. BUT: classification probably still makes more sense, because a regression framework assumes equal change  in the output variable for the same amount of change in the input variable suppose we have a place that's very unlikely to gain erg (i.e. the prediction
is close to 0), and we increase some input value that makes gain more likely(e.g. we increase the amount of nearby erg, or upwind erg, or ...) then we expect a big change in the output variable (from say 0.01 to 0.5)
but if we have a place that's *likely* to gain erg (the prediction is already
close to 1), then it doesn't make sense to increase the prediction by the
same amount, because a prediction >1 doesn't really make  sense
this is why we usually use logistic regression rather than linearregression for binary responseswhich is a long-winded way of saying we should probably do classification instead
```{r}
train <- subset(ss2, sample == TRUE)
test <- subset(ss2, sample == FALSE)
```
random forest model
```{r}
if (file.exists("rf.RData")) {
    load("rf.RData")
} else {
    rf <- randomForest(formula= factor(change2) ~ . - x - y ,
                       data = train, n.trees=250,interaction.depth=7, do.trace=1,
                       type="classification", proximity=TRUE)
    save("rf",file="rf.RData")
}

rf_tune=tuneRF(formula= factor(change2) ~ . - x - y ,
             data = train, n.trees=c(0,75,200,500),interaction.depth=c(5,6,7,8), do.trace=1,
             type="classification", proximity=TRUE)
```
random forests plots
```{r}
plot(rf)
plot(rf$predicted)
```
predicting with the test data set
```{r}
pred <- predict(rf, newdata=test,type="class")
table(pred,test$change2)
```
confusionMatrix
```{r}
ff <- function(x) factor(x,levels=0:3,labels=c("no gain","gain","loss","no loss"))
conf500= caret::confusionMatrix(ff(pred),ff(test$change2))
save("conf500",  file="saved_conf-500.RData")
```
fit random forest model vy two methods
```{r}
rf_fit <- train(as.factor(change2) ~ ., 
                data = train, 
                method = "ranger")
rf_fit2 <- ranger(change2 ~ ., data = train, num.trees = 500, mtry = 6, importance = "impurity", min.node.size = 3, replace = TRUE, num.threads = 3)
save("rf_fit",  file="saved_rf_fit.RData")
```
predicting fit random forest model with the test data set
```{r}
rf_pred <- predict(rf_fit, dat_upsample)
confusionMatrix(rf_pred, as.factor(dat_upsample$change2))
```
rf$votes is the number of votes for each category
for each data point, calculate the proportion of trees that got the right answer
```{r}
mm <- match(train$change,0:3) ## which column matches?
## this gives us the column number we want in each row
correct_prop <- rf$votes[cbind(seq(nrow(rf$votes)), mm)]
any(is.na(correct_prop)) ## no missing values
any(is.na(cr(correct_prop))) ## no missing values
rr <- rgb(cr(correct_prop)/255) ## no missing values
any(is.na(rr))
testpoint <- SpatialPointsDataFrame(cbind(train$x, train$y), train)
plot(testpoint$x,testpoint$y,col=rgb(cr(correct_prop)/255),
     pch=16) ##,pch=".",cex=3)
points(test$x,test$y,pch=".")

head(rf$votes)
head(a2$change)
npts <- nrow(rf$votes)
```
RMSE (Root Mean Square Error)of this optimal random forest
```{r}
sqrt(rf$mse[which.min(rf$mse)])
testpoint <- SpatialPointsDataFrame(cbind(a2$x, a2$y), a2)
plot(testpoint$x,testpoint$y,col=rgb(cr(res)/255),
     pch=16) ##,pch=".",cex=3)
```